# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
setwd("C:/Users/HomeUser/Documents/Spécialité Data Science/Reproducible Research/repos")

data <- read.csv("./activity.csv")
data$date <- as.Date(data$date,"%Y-%m-%d")

head(data)
summary(data)
```
## What is mean total number of steps taken per day?
```{r echo=TRUE}
options(scipen = 1, digits = 2) #get rid of exponential format

aggdata <- tapply(data$steps,data$date, sum, na.rm = T)
barplot(aggdata, names.arg = as.Date(names(aggdata)) 
     , main="Histogram of the total number of steps taken per day"
     ,ylab="Steps", xlab="Day"
     )

meansteps <- round(mean(aggdata))
mediansteps <- round(median(aggdata))

```
Mean total number of steps taken per day: `r meansteps`.

Median total number of steps taken per day: `r mediansteps`.

## What is the average daily activity pattern?
```{r echo=TRUE}
dailydata <- tapply(data$steps,data$interval, mean, na.rm = T)
plot(as.numeric(names(dailydata)),dailydata
     ,type='l'
     ,main="Average daily activity"
     ,xlab="5-minute interval"
     ,ylab="average number of steps per day"
     )
indmaxinterv = which(dailydata == max(dailydata))
maxinterv = names(dailydata[indmaxinterv])

```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? `r maxinterv`

## Imputing missing values
```{r echo=TRUE}
nbna = sum(is.na(data))
print("Total number of missing values in the dataset: ")
print(nbna)

#Strategy: Use of the mean for that 5-minute interval, calculated early in dailydata
newdata <- data
indexes_na <- which(is.na(data$step))
newdata[indexes_na,]$steps <- sapply(data[indexes_na,]$interval, function(x){
        dailydata[as.character(x)][[1]]
})

#check: first lines of newdata should contain average values corresponding to the interval considered
head(newdata)
dailydata[c(as.character(seq(0,25,by=5)))]

#Plot histogram
aggdata <- tapply(newdata$steps,newdata$date, sum, na.rm = T)
barplot(aggdata, names.arg = as.Date(names(aggdata)) 
     , main="Histogram of the total number of steps taken per day"
     ,ylab="Steps", xlab="Day"
     )

meansteps <- round(mean(aggdata))
mediansteps <- round(median(aggdata))

```
Mean total number of steps taken per day: `r meansteps`.

Median total number of steps taken per day: `r mediansteps`.

Imputing missing values raises overall number of steps!

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
Sys.setlocale(category = "LC_ALL", locale = "C")

newdata$week <- ifelse(weekdays(newdata$date) %in% c("Sunday","Saturday"),"weekend","weekday")
newdata$week <- factor(newdata$week)

par(mfrow = c(2,1))

newdailydataWE <- tapply(newdata[newdata$week == "weekend",]$steps,newdata[newdata$week == "weekend",]$interval, mean, na.rm = T)
plot(as.numeric(names(newdailydataWE)),newdailydataWE
     ,type='l'
     ,main="weekend"
     ,xlab="interval"
     ,ylab="Number of steps"
     )

newdailydataWD <- tapply(newdata[newdata$week == "weekday",]$steps,newdata[newdata$week == "weekday",]$interval, mean, na.rm = T)
plot(as.numeric(names(newdailydataWD)),newdailydataWD
     ,type='l'
     ,main="weekday"
     ,xlab="interval"
     ,ylab="Number of steps"
     )

Sys.setlocale(category = "LC_ALL", locale = "")
```